Hybrid platforms and bargaining power

### 2024-01-05 src/paper/application.tex
Reorganize paper

In the benchmark model, the platform will be able to set and commit to any lump-sum entry fee it wishes to.\footnote{
    The focus of this paper is on fees paid and entry decisions made by fringe producers, but I will also consider the case when consumers might also face entry fees or subsidies.
}
The main contribution of this paper is to deviate from this usual assumption, and instead consider a setting in which the platform and the fringe engage in some form of bargaining over the total profits.
To model this in a theoretically sound, but still tractable way, I borrow a solution concept from cooperative game theory.
Similarly to a number of other papers in the industrial organization literature \parencite[e.g.][]{montez2007downstream,hart1990property,levy1997individual,inderst2003bargaining,brugemann2019intra}, I assume that the resulting allocation for each player is given by their Shapley value.
This value (and the game itself) depends on who is assumed to participate in the bargaining process.
I will consider two different cases.

\paragraph{One-sided case} In the first setting, the platform and the fringe firms bargain over the total profits of the platform and the fringe taken together, and the consumers are not included in the negotiations.
The following assumption describes this case.
\begin{assumption}
    \label{ass:profit_sharing}
    Consider the following cooperative game $\mathcal{G} = (\mathcal{N}, v)$.
    The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i\}, i \in [0, N_F]$.
    The value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
    \begin{align*}
        v(S) = \begin{cases}
            0 & \text{if } P \notin S \\
            N_C f(N_P, n_F(S)) & \text{if } P \in S
        \end{cases},
    \end{align*}
    where $n_F(S)$ is the measure of fringe firms in $S$.
    Each type of player gets their Shapley value in this game:
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= \varphi_P(\mathcal{G}), \\
        \pi_F(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}),
    \end{align*}
    where $\varphi_i(\mathcal{G})$ denotes the Shapley value of player(s) $i$ in game $\mathcal{G}$.
\end{assumption}

While this assumption might seem somewhat abstract, the resulting profit allocations are remarkably simple, and can be expressed as integrals (expectations) over the marginal contributions of the players.
\begin{proposition}
    \label{prop:profit_sharing}
    Consider the case when only the platform and the fringe firms participate in the bargaining process.
    Then, the resulting profit shares are given by
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= N_C \int_0^1 f(N_P, s N_F) \ds, \\
        \pi_F(N_P, N_F, N_C) &= N_C \int_0^1 s N_F \partial_2 f(N_P, s N_F) \ds.
    \end{align*}
\end{proposition}

\begin{proof}[proof of \cref{prop:profit_sharing}]
    This result is a direct application of Proposition 3 and Corollary 2 in \textcite{stancsics2023value}.
\end{proof}

As mentioned before, one main justification for this profit allocation rule its intuitive behavior in terms of comparative statistics.
To illustrate this, let us examine what happens when one varies the substitutability between the fringe firms.
As it turns out, the platform's share increases if the fringe firms are more substitutable.
The following observation demonstrates this idea.
\begin{observation}
    \label{prop:outcome_based_bargaining_power}
    Fix some $N_P, N_C \geq 0$. Let $f, \tilde{f}: \mathbb{R}^+_0 \times \mathbb{R}^+_0 \to \mathbb{R}^+_0$ two different profit functions such that $f(N_P, N_F) = \tilde{f}(N_P, N_F)$ for some $N_P, N_F$ and $f(N_P, n_F) \leq \tilde{f}(N_P, n_F)$ for all $n_F < N_F$.
    Furthermore, let us denote the corresponding platform profit shares by $\pi_P$ and $\tilde{\pi}_P$.

    Then, $\pi_P \leq \tilde{\pi}_P$.
\end{observation}

@@

\paragraph{Two-sided case} Alternatively, in certain settings, it might be appropriate to assume that the consumers (or, more generally, the entities on the other side of the market) also engage in the bargaining process.
The cooperative game describing this situation is formalized in \cref{ass:profit_sharing_two_sided}.
\begin{assumption}
    \label{ass:profit_sharing_two_sided}
    Consider the following cooperative game $\mathcal{G} = (\mathcal{N}, v)$.
    The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i, C_j\}, i \in [0, N_F], j \in [0, N_C]$.
    The value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
    \begin{align*}
        v(S) = \begin{cases}
            0 & \text{if } P \notin S \\
            n_C(S) f(N_P, n_F(S)) & \text{if } P \in S
        \end{cases},
    \end{align*}
    where $n_F(S)$ and $n_C(S)$ is the measure of fringe firms and consumers, respectively, in $S$.
    Each type of player gets their Shapley value in this game:
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= \varphi_P(\mathcal{G}), \\
        \pi_F(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}), \\
        \pi_C(N_P, N_F, N_C) &= \varphi_F(\mathcal{G}),
    \end{align*}
    where $\varphi_i(\mathcal{G})$ denotes the Shapley value of player(s) $i$ in game $\mathcal{G}$.
\end{assumption}

As in the one-sided case, simple expressions exist for these various shares.
\begin{proposition}
    \label{prop:profit_sharing_two_sided}
    Consider the case when only the platform and the fringe firms participate in the bargaining process.
    Then, the resulting profit shares are given by
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= N_C \int_0^1 s f(N_P, s N_F) \ds, \\
        \pi_F(N_P, N_F, N_C) &= N_C \int_0^1 s^2 N_F \partial_2 f(N_P, s N_F) \ds \\
        \pi_C(N_P, N_F, N_C) &= N_C \int_0^1 s f(N_P, s N_F) \ds.
    \end{align*}
\end{proposition}
\begin{proof}[proof of \cref{prop:profit_sharing}]
    This result is a direct application of Proposition 11 in \textcite{stancsics2023value}.
\end{proof}
One thing to note is that the platform's and fringe firms' shares are lower than in the one-sided case.
This is due to the fact of needing to share the pie with the consumers, too.
Additionally, the platform's share is identical to the consumers' (aggregated) share.\footnote{
    This stems from the linearity of the profit function in the number of consumers.
}
This foreshadows the idea that -- even though it might not be welfare-maximizing -- what's good for the platform might also good for the consumers in the two-sided case.

@@

Finally, in order to guarantee a unique equilibrium, let us make the following, an additional assumption about the profit function.
\begin{assumption}
    \label{ass:single_crossing}
    Let $f$ be such that the following holds for all $N_F, N_P, N_C \geq 0$:
    \begin{align*}
        \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_F} < 0 \text{ or } \frac{\partial^2 \pi_F(N_P, N_F, N_C)}{\partial N_F^2} < 0
    \end{align*}
\end{assumption}
This assumption essentially guarantees that the profit of the fringe (as a function of the number of entrants) has at most a single crossing with the --linear -- total entry cost (apart from the obvious $N_F=0$ intersection).
The interpretation of the assumption itself is simple: total fringe profits, as a function of the number of fringe firms, must be either concave or hump-shaped.\footnote{
    As it turns out, even though total industry profits are assumed to be increasing in $n_f$, it is not necessarily true for fringe profits.
    For certain profit functions, it happens that even though the size of the pie increases, the bargaining position of the fringe deteriorates so much that they get a smaller share even in absolute value.
}
What it requires of $f$, however, is not so obvious, and also depends on whether one assumes one-sided or two-sided bargaining.
Nevertheless, when microfounding the profit function, one can simply check whether this assumption holds or not.
In particular, it will be satisfied for the logit-type demand in \cref{sec:example}.

@@

For the following two propositions, let us ignore free entry and equilibrium behavior, and consider the number of fringe entrants $N_F$ as fixed.
The following statement claims that -- in both the one-sided and two-sided bargaining cases -- the platform's profits are increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Assume that $f$ is continuously differentiable with respect to $N_P$ and also twice differentiable.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0.
    \end{align*}
\end{proposition}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the slice of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is only guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.\footnote{
    In fact, one can show that this is the case if the cross-derivatives of $f$ are negative.
}
As this result shows, this decrease cannot be large enough to decrease the platform's profits in absolute terms.
\cref{fig:increase_N_P_platform} shows an example of this situation.

@@

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has a specific, additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.\footnote{
    Note that it does not mean that their products are substitutes in the usual, demand theory sense.
    They still might be complements to each other.
}
\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let $f$ have the following, additive form:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$.
    Also assume that $g$ is twice differentiable and that $g'' < 0$.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
    Furthermore,
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_P} < 0.
    \end{align*}
\end{proposition}

@@

In the main model, I consider the case when the total profit function only includes the sales of the platform's and fringe firms' products.
That is, the platform does not charge any entry fees to the consumers.
Here, I consider two variants of the model: a benchmark where the platform can unilaterally commit to any entry fee, and a bargaining model where the platform and the fringe firms negotiate over the distribution of profits.

As an extension, I also examine what happens if the platform is able to extract consumer surplus by charging a flat entry fee, and the platform and the firms bargain over this larger sum.
Furthermore, I also consider the case of two-sided bargaining, where the consumers also participate in the negotiation over the value produced.

@@

The participants negotiate over the aggregate profits that they expect to obtain in the subsequent period (assuming the same non-collusive, monopolistic pricing, as before).
I assume that bargaining outcomes are determined according to \cref{ass:profit_sharing}, where $\Pi(N_P, N_F, N_C)$ is defined as the sum of profits that the platform and the entrants acquire by selling their products.
From \cref{eq:optimal_profit}, the total profit achieved by the platform and $N_F$ fringe firms is given by
\begin{align*}
    \Pi(N_P, N_F, N_C) = N_C \underbracket{\mu \frac{n V_F + N_P V_P}{n V_F + N_P V_P + 1}}_{\coloneqq f(N_P, N_F)}.
\end{align*}
Based on \cref{prop:profit_sharing}, one can even obtain closed-form expressions for platform and fringe profit shares for this class of demand systems.
\begin{proposition}
    \label{prop:platform_profits_bargaining}
    The platform's total profits are
    \begin{align*}
        \pi^t_P &= N_C \int_0^1 f(N_P, sN_F) \ds \\
                &= \mu N_C \left[ 1 - \frac{\log \left(1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} \right].
    \end{align*}
    The total profits of the whole fringe (including investment costs) are given by
    \begin{align*}
        \pi^t_F = \mu N_C \left[ \frac{\log \left( 1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} - \frac{1}{N_P V_P + N_F V_F + 1} \right] - I_F N_F .
    \end{align*}
\end{proposition}

@@

@@

@@

% TODO: do I want a section on Shapley-values, or move some cooperative stuff here?

@@

\begin{lemma}
    \label{lem:single_crossing_affine_transformation}
    Let $g: \mathbb{R}^+ \to \mathbb{R}$ be a twice continuously differentiable function.
    Define
    \begin{align*}
        G(x) = \int_0^x g(sx) \ds.
    \end{align*}
    Assume that for any $x \geq 0$, $G'(x) < 0$ or $G''(x) < 0$.

@@

\begin{proof}[Proof of \cref{prop:share_of_platform}]
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F, N_C)}{\partial N_P} &= N_C \frac{\partial}{\partial N_P} \int_0^1 w(s) f(N_P, N_F) \ds \\
        &= \int_0^1 w(s) \underbrace{\frac{\partial f(N_P, sN_F)}{\partial N_P}}_{> 0} \ds > 0
    \end{align*}
    for any continuous $w(s)$.
    In particular, this holds for $w(s) = 1$ and $w(s) = s$, which correspond to assumptions \cref{ass:profit_sharing} and \cref{ass:profit_sharing_two_sided}, respectively.
\end{proof}

\begin{proof}[Proof of \cref{prop:share_of_fringe}]
    Remember that
    \begin{align*}
        \pi_F(N_P, N_F, N_C) = N_C \int_0^1 w(s) N_F \partial_2 f(N_P, s N_F) \ds,
    \end{align*}
    where $w(s)$ is either $s$ or $s^2$, depending on whether one assumes one or two-sided bargaining.
    By assumption, $f$ is twice continuously differentiable, therefore $\partial_2$ is also continuously differentiable in $N_P$.
    Thus the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} &= \frac{\partial}{\partial N_P} N_C \int_0^1 w(s) N_F \partial_2 f(N_P, s N_F) \ds \\
        &= N_C \int_0^1 w(s) N_F \frac{\partial}{\partial N_P} \partial_2 f(N_P, s N_F) \ds \\
        &= N_C \int_0^1 w(s) N_F \partial^2_{12} f(N_P, s N_F) \ds.
    \end{align*}

    As $w(s) > 0$ for $s > 0$, if $\partial_2 f(N_P, s N_F)$ has the same sign over $[0, N_F]$, then the integral also has the same sign.
    Formally,
    \begin{align*}
        \forall n_F \in [0, N_F] \enspace \frac{\partial^2 f(N_P, n_F)}{\partial N_P \partial n_F} < 0 &\implies \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} < 0 \\
        \forall n_F \in [0, N_F] \enspace \frac{\partial^2 f(N_P, n_F)}{\partial N_P \partial n_F} > 0 &\implies \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} > 0
    \end{align*}
\end{proof}

>>>>

In the benchmark model, the platform will be able to set and commit to any lump-sum entry fee it wishes to.\footnote{
    The focus of this paper is on fees paid and entry decisions made by fringe producers, but I will also consider the case when consumers also face entry fees or subsidies.
}
The main contribution of this paper is to deviate from this usual assumption, and instead consider a setting in which the platform and the fringe engage in some form of bargaining over the total profits.
To model this in a tractable way, I propose a profit division rule that is based on the Shapley value \parencite{shapley1953additive}, or, more generally, random order values \parencite{weber1988probabilistic}.
This way of modelling bargaining outcomes has many precedents in the industrial organization literature \parencite[e.g.][]{montez2007downstream,hart1990property,levy1997individual,inderst2003bargaining,brugemann2019intra}.
In the main text, I simply propose a formula for the profit shares without going into the details of the underlying cooperative game.
The interested reader can find a more detailed discussion of how it relates to random order values in \cref{sec:cooperative_game}.

\begin{assumption}
    \label{ass:profit_sharing}
    Let $N_P, N_F, N_C \geq 0$ be the number of platform products, fringe firms and consumers, respectively.
    Thus, the value that players create together is given by $\Pi(N_P, N_F, N_C) = N_C f(N_P, N_F)$.
    Let us denote the platform's and the fringe's profit shares by $\pi_P(N_P, N_F, N_C)$ and $\pi_F(N_P, N_F, N_C)$, respectively.
    Then, the following holds:
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= N_C \int_0^1 w_P(s) f(N_P, s N_F) \ds, \\
        \pi_F(N_P, N_F, N_C) &= N_C \int_0^1 w_F(s) N_F \partial_2 f(N_P, s N_F) \ds.
    \end{align*}
    Furthermore, $w_P \geq 0$ and $w_F \geq 0$ are such that $\pi_P(N_P, N_F, N_C) + \pi_F(N_P, N_F, N_C) \leq \Pi(N_P, N_F, N_C)$.
\end{assumption}

Although the reason behind choosing this particular form for profit shares is discussed in detail in the appendix, let me highlight some important observations.
First, player's profit shares are closely related to their (weighted average) marginal contributions to the total value.
In the case of the platform, its contribution is the total value, as it is indispensable, while the fringe firms' contribution is captured by the partial derivative of the total value with respect to the number of fringe firms.
Second, the platform and the fringe do not necessarily get the entire value they create.
This is due to the fact that I also want to allow for the possibility of consumers participating in the bargaining process and getting a share of the profits, too
Finally, the profit shares depend on two types of factors: the production function $f$ and $w_P$ and $w_F$, which I will call bargaining weights.
The former is determined by the technology and demand structure, while the latter can be thought of as some king of intrinsic bargaining power.

In the later sections, I will consider specific examples for $f$ and $w_P$ and $w_F$, motivated by a Logit-like demand system, and (weighted) Shapley values, respectively.
However, for a number of results, this level of detail is not necessary, and these quite general assumptions are sufficient.

As mentioned before, one main justification for this profit allocation rule its intuitive behavior in terms of comparative statistics.
To illustrate this, let us examine what happens when one varies the substitutability between the fringe firms.
As it turns out, the platform's share increases if the fringe firms are more substitutable.
The following observation demonstrates this idea.
\begin{proposition}
    \label{prop:outcome_based_bargaining_power}
    Fix some $N_P, N_C \geq 0$. Let $f, \tilde{f}: \mathbb{R}^+_0 \times \mathbb{R}^+_0 \to \mathbb{R}^+_0$ two different profit functions such that $f(N_P, N_F) = \tilde{f}(N_P, N_F)$ for some $N_P, N_F$ and $f(N_P, n_F) \leq \tilde{f}(N_P, n_F)$ for all $n_F < N_F$.
    Furthermore, let us denote the corresponding platform profit shares by $\pi_P$ and $\tilde{\pi}_P$.

    Then, $\pi_P \leq \tilde{\pi}_P$.
\end{proposition}

@@

@@

Finally, in order to guarantee a unique equilibrium, let us make the following, an additional assumption about the profit function.
\begin{assumption}
    \label{ass:single_crossing}
    Let $f$ be such that the following holds for all $N_F, N_P, N_C \geq 0$:
    \begin{align*}
        \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_F} < 0 \text{ or } \frac{\partial^2 \pi_F(N_P, N_F, N_C)}{\partial N_F^2} < 0
    \end{align*}
\end{assumption}
This assumption essentially guarantees that the profit of the fringe (as a function of the number of entrants) has at most a single crossing with the --linear -- total entry cost (apart from the obvious $N_F=0$ intersection).
The interpretation of the assumption itself is simple: total fringe profits, as a function of the number of fringe firms, must be either concave or hump-shaped.\footnote{
    As it turns out, even though total industry profits are assumed to be increasing in $n_f$, it is not necessarily true for fringe profits.
    For certain profit functions, it happens that even though the size of the pie increases, the bargaining position of the fringe deteriorates so much that they get a smaller share even in absolute value.
}
What it requires of $f$, however, is not so obvious, and also depends on what one assumes about the innate bargaining weights $w_F$ and $w_P$.
Nevertheless, when microfounding the profit function, one can simply check whether this assumption holds or not.
In particular, it will be satisfied for the logit-type demand in \cref{sec:example}.
% TODO: conjecture: it only has to hold for $f(N_P, N_F)$.

@@

For the following two propositions, let us ignore free entry and equilibrium behavior, and consider the number of fringe entrants $N_F$ as fixed.
The following statement claims that the platform's profits are increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Assume that $f$ is continuously differentiable with respect to $N_P$ and also twice differentiable.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0.
    \end{align*}
\end{proposition}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the slice of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is only guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.\footnote{
    In fact, one can show that this is the case if the cross-derivatives of $f$ are negative.
}
As this result shows, this decrease cannot be large enough to decrease the platform's profits in absolute terms.
\cref{fig:increase_N_P_platform} shows an example of this situation.

@@

Finally, let us conclude this section by establishing a stronger result for a more restrictive class of profit functions.
In the following, I assume that the profit function has a specific, additive form.
Intuitively, this means that -- in terms of total profits generated -- the platform's and the fringe firms' products are substitutable to each other according to some constant ratio.
\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let $f$ have the following, additive form:
    \begin{align*}
        f(N_P, N_F) = g(\alpha N_P + \beta N_F)
    \end{align*}
    where $\alpha, \beta > 0$.
    Also assume that $g$ is twice differentiable and that $g'' < 0$.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
    Furthermore,
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_P} < 0.
    \end{align*}
\end{proposition}

@@

In the main model, I consider the case when the total profit function only includes the sales of the platform's and fringe firms' products.
That is, the platform does not charge any entry fees to the consumers.
Here, I consider two variants of the model: a benchmark where the platform can unilaterally commit to any entry fee, and a bargaining model where the platform and the fringe firms negotiate over the distribution of profits.
Furthermore, I focus on a particular set of weight functions $w_P$ and $w_F$, which correspont to the Shapley-values of the players in the case of one-sided bargaining.

As an extension, I also examine what happens when the platform and the fringe firms have unequal bargaining weights, while still retaining more structure than the very general bargaining rule from \cref{ass:profit_sharing}.
In particular, I look at the case of bargaining outcomes being describet by weighted values, with all fringe players sharing the same weight.
Furthermore, I also consider the case of two-sided bargaining, where the consumers also participate in the negotiation over the value produced.

@@

The participants negotiate over the aggregate profits that they expect to obtain in the subsequent period (assuming the same non-collusive, monopolistic pricing, as before).
I assume that bargaining outcomes are determined according to \cref{ass:profit_sharing}, where $\Pi(N_P, N_F, N_C)$ is defined as the sum of profits that the platform and the entrants acquire by selling their products.
From \cref{eq:optimal_profit}, the total profit achieved by the platform and $N_F$ fringe firms is given by
\begin{align*}
    \Pi(N_P, N_F, N_C) = N_C \underbracket{\mu \frac{n V_F + N_P V_P}{n V_F + N_P V_P + 1}}_{\coloneqq f(N_P, N_F)}.
\end{align*}

Let us assume that the bargaining results in profits being shared according to players' Shapley values.
These Shapley values satisfy \cref{ass:profit_sharing}, with $w_P(s) \equiv 1$ and $w_F(s) = s$.\footnote{
    For a discussion and proof of this claim, see \cref{sec:cooperative_game}.
}
Based on these, one can even obtain closed-form expressions for platform and fringe profit shares for this class of demand systems.
\begin{proposition}
    \label{prop:platform_profits_bargaining}
    The platform's total profits are
    \begin{align*}
        \pi^t_P &= N_C \int_0^1 f(N_P, sN_F) \ds \\
                &= \mu N_C \left[ 1 - \frac{\log \left(1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} \right].
    \end{align*}
    The total profits of the whole fringe (including investment costs) are given by
    \begin{align*}
        \pi^t_F = \mu N_C \left[ \frac{\log \left( 1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} - \frac{1}{N_P V_P + N_F V_F + 1} \right] - I_F N_F .
    \end{align*}
\end{proposition}

@@

In the previous sections, I assumed that the platform and the fringe firms split total profits according to their Shapley values.
Because of the symmetry property of the Shapley value, this excludes the possibility of players having some kind of innate bargaining power, which is not related to the profit functions.
In this extension, I will relax this assumption, and assume instead that profits are shared according to \emph{weighed} Shapley values.
These are a generalization of the Shapley value, where each player has a weight $w_i$ that can, in certain settings, be thought of as a parameter describing bargaining power.\footnote{
    \textcite{hart1996bargaining} and \textcite{stole1996intra} provide foundations for this interpretation.
}
Similarly to the standard Shapley value, one can show that outcomes based on weighted values also satisfy \cref{ass:profit_sharing}.
Specifically, when the weight of the platform is $\lambda_P$, the functions $w_P$ and $w_F$ are given by
\begin{align*}
    w_P(s) &= \lambda_P s ^ {\lambda_P - 1} \\
    w_F(s) &= s ^ {\lambda_P}.
\end{align*}

TODO: can I show that \cref{ass:single_crossing} still holds for any $\lambda_P$? If not, focus on a specific case, such as the one on the graphs.

@@

\section{The platform game and the (weighted) Shapley value}
\label{sec:cooperative_game}

In this section, I describe how the profit sharing rule from \cref{sec:profit_sharing} can be derived from various cooperative games.
In each of the following subsections, I formally define a cooperative game that models the platform setting, and derive the (weighted) Shapley values of the various players.
I show that they correspond to the formulas given in \cref{sec:profit_sharing}, with a specific choice of weight functions $w_P$ and $w_F$ for each case.

I start with the simplest case: one-sided bargaining and the usual Shapley value \parencite{shapley1953additive}.
Then, I look at weighted values \parencite{weber1988probabilistic} in the same cooperative game.
Finally, I consider the case when the consumers also participate in the bargaining process, and derive the corresponding Shapley values.

\subsection{One-sided case}
\label{sec:cooperative_game_one_sided}

In the first setting, the platform and the fringe firms bargain over the total profits they generate together, and the consumers are not included in the negotiations.
The following cooperative game describes this case.

Consider the game $\mathcal{G} = (\mathcal{N}, v)$, characterized by the set of players $\mathcal{N}$ and its characteristic function $v$.
The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i\}, i \in [0, N_F]$.
Because the platform is indispensible, the value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
\begin{align*}
    v(S) = \begin{cases}
        0 & \text{if } P \notin S \\
        N_C f(N_P, n_F(S)) & \text{if } P \in S
    \end{cases},
\end{align*}
where $n_F(S)$ is the measure of fringe firms in $S$.

The following proposition describes the Shapley value of the platform ($\varphi_P(\mathcal{G})$) and the fringe firms ($\varphi_F(\mathcal{G})$) in this game.

\begin{proposition}
    \label{prop:profit_sharing_one_sided}
    Consider the cooperative game above.
    Then, the Shapley values of the platform and the fringe firms are given by
    \begin{align*}
        \varphi_P(\mathcal{G}) &= N_C \int_0^1 f(N_P, s N_F) \ds, \\
        \varphi_P(\mathcal{F}) &= N_C \int_0^1 s N_F \partial_2 f(N_P, s N_F) \ds.
    \end{align*}
\end{proposition}

\begin{proof}[proof of \cref{prop:profit_sharing_one_sided}]
    This result is a direct application of Proposition 3 and Corollary 2 in \textcite{stancsics2023value}.
\end{proof}

As an immediate consequence, if bargaining outcomes are described by the Shapley value, then the resulting allocations satisfy \cref{ass:profit_sharing} with $w_P(s) \equiv 1$ and $w_F(s) = s$.

\subsection{Weighted value}
\label{sec:cooperative_game_weighted}

Now let us generalize the previous result, and assume instead that players' shares are described by their weighted values,
As before, let us start by formally describing the cooperative game.
The set of players and the characteristic function are the same as in \cref{sec:cooperative_game_one_sided}, but now the game is additionally endowed with a weight system $\mathbf{\lambda} = \{\lambda_P, \lambda_{F}(i)\}$.
Let us assume that all fringe players have the same weight $\lambda_{F}(i) \equiv 1$.

Then, the weighted values of the players are described by the following proposition.

\begin{proposition}
    \label{prop:profit_sharing_weighted}
    Consider the cooperative game above.
    Then, the weighted values of the platform and the fringe firms are given by
    \begin{align*}
        \varphi_P(\mathcal{G}) &= N_C \int_0^1 \lambda_P s ^ {\lambda_P - 1} f(N_P, s N_F) \ds, \\
        \varphi_F(\mathcal{G}) &= N_C \int_0^1 s ^ {\lambda_P} N_F \partial_2 f(N_P, s N_F) \ds.
    \end{align*}
\end{proposition}

\begin{proof}[proof of \cref{prop:profit_sharing_weighted}]
    This result is a direct application of Proposition 4 in \textcite{stancsics2023value}.
\end{proof}

As before, the resulting allocations satisfy \cref{ass:profit_sharing} with $w_P(s) \equiv \lambda_P s ^ {\lambda_P - 1}$ and $w_F(s) = s ^ {\lambda_P}$.

\subsection{Two-sided case}
\label{sec:cooperative_game_two_sided}

Finally, in certain settings, it might be appropriate to assume that the consumers (or, more generally, the entities on the other side of the market) also engage in the bargaining process.
The cooperative game describing can be formalized as follows.

The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i, C_j\}, i \in [0, N_F], j \in [0, N_C]$.
The value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
\begin{align*}
    v(S) = \begin{cases}
        0 & \text{if } P \notin S \\
        n_C(S) f(N_P, n_F(S)) & \text{if } P \in S
    \end{cases},
\end{align*}
where $n_F(S)$ and $n_C(S)$ is the measure of fringe firms and consumers, respectively, in $S$.

As in the one-sided case, simple expressions exist for the Shapley value of the platform ($\varphi_P(\mathcal{G})$), fringe firms ($\varphi_F(\mathcal{G})$) and consumers ($\varphi_C(\mathcal{G})$).

\begin{proposition}
    \label{prop:profit_sharing_two_sided}
    Consider the case when only the platform and the fringe firms participate in the bargaining process.
    Then, the resulting profit shares are given by
    \begin{align*}
        \varphi_P(\mathcal{G}) &= N_C \int_0^1 s f(N_P, s N_F) \ds, \\
        \varphi_F(\mathcal{G}) &= N_C \int_0^1 s^2 N_F \partial_2 f(N_P, s N_F) \ds \\
        \varphi_C(\mathcal{G}) &= N_C \int_0^1 s f(N_P, s N_F) \ds.
    \end{align*}
\end{proposition}
\begin{proof}[proof of \cref{prop:profit_sharing_two_sided}]
    This result is a direct application of Proposition 11 in \textcite{stancsics2023value}.
\end{proof}

One thing to note is that the platform's and fringe firms' shares are lower than in the one-sided case.
This is due to the fact of needing to share the pie with the consumers, too.
Additionally, the platform's share is identical to the consumers' (aggregated) share.\footnote{
    This stems from the linearity of the profit function in the number of consumers.
}
This foreshadows the idea that -- even though it might not be welfare-maximizing -- what's good for the platform might also good for the consumers in the two-sided case.
Finally, as in both examples before, these values satisfy \cref{ass:profit_sharing} with $w_P(s) = s$ and $w_F(s) = s^2$.

@@

@@

\begin{lemma}
    \label{lem:single_crossing_affine_transformation}
    Let $g: \mathbb{R}^+ \to \mathbb{R}$ be a twice continuously differentiable function.
    Define
    \begin{align*}
        G(x) = \int_0^1 g(sx) \ds.
    \end{align*}
    Assume that for any $x \geq 0$, $G'(x) < 0$ or $G''(x) < 0$.

@@

\begin{proof}[Proof of \cref{prop:share_of_platform}]
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F, N_C)}{\partial N_P} &= N_C \frac{\partial}{\partial N_P} \int_0^1 w_P(s) f(N_P, N_F) \ds \\
        &= \int_0^1 w_P(s) \underbrace{\frac{\partial f(N_P, sN_F)}{\partial N_P}}_{> 0} \ds > 0
    \end{align*}
    for any non-negative, continuous $w_P(s)$.
\end{proof}

\begin{proof}[Proof of \cref{prop:share_of_fringe}]
    Remember that
    \begin{align*}
        \pi_F(N_P, N_F, N_C) = N_C \int_0^1 w_F(s) N_F \partial_2 f(N_P, s N_F) \ds,
    \end{align*}
    By assumption, $f$ is twice continuously differentiable, therefore $\partial_2$ is also continuously differentiable in $N_P$.
    Thus, the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} &= \frac{\partial}{\partial N_P} N_C \int_0^1 w_F(s) N_F \partial_2 f(N_P, s N_F) \ds \\
        &= N_C \int_0^1 w_F(s) N_F \frac{\partial}{\partial N_P} \partial_2 f(N_P, s N_F) \ds \\
        &= N_C \int_0^1 w_F(s) N_F \partial^2_{12} f(N_P, s N_F) \ds.
    \end{align*}

    As $w_F(s) \geq 0$ for $s > 0$, if $\partial_2 f(N_P, s N_F)$ has the same sign over $[0, N_F]$, then the integral also has the same sign.
    Formally,
    \begin{align*}
        \forall n_F \in [0, N_F] \enspace \frac{\partial^2 f(N_P, n_F)}{\partial N_P \partial n_F} < 0 &\implies \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} < 0 \\
        \forall n_F \in [0, N_F] \enspace \frac{\partial^2 f(N_P, n_F)}{\partial N_P \partial n_F} > 0 &\implies \frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P} > 0
    \end{align*}
\end{proof}
